### 2024-08-19 paper.tex
reorganize 8

\section{Finding the irreducible components of \(U_0\)}
Let \(X\) be nilpotent with Jordan basis \((e_{ij}')_{1 \leq i \leq k, 1 \leq j \leq \lambda_i'}\).
Let \(V = \{(A_{a,b}, U) : \forall i. \; A_{a,b} U_{i + 1} \subseteq U\}\).
Let
\[U_0 = \{(A_{0,b}, U) \in V\}.\]
We will find the irreducible components of \(U_0\).
\par We write \(b_{ij}\) to denote the projection of \(b \in \mathbb{C}^m\) onto \(e_{ij}\).
For each row \(i\), let \(p_i(b) = \max\{j : b_{ij} \neq 0\}\) (the maximum of the empty set is zero).
Then set \(q_i(b) = \lambda_i - p_i(b)\).
When it is clear enough from context, we will just write \(p_i\) and \(q_i\) instead of \(p_i(b)\) and \(q_i(b)\).
\par For any \(I \subseteq \{1, ..., k\}\) and any values \((\rho_i)_{i \in I}\) %such that (1) \(\rho_i \leq \lambda_i\), (2) \(\rho_i\) is decreasing with \(i\), and (3) \(\lambda_i - \rho_i\) is decreasing with \(i\)
, define 
\[B_{I, (\rho_i)} := \{b \in \mathbb{C}^m : \]
\[\{(p_i, q_i) : i \in I\} = \{(p_i, q_i) : n > q_i \land p_i = \max_{j : q_j = q_i} p_j > \max_{j : q_j < q_i} p_j \} \] 
\[\land \forall i \in I. \; p_i = \rho_i \}.\]
Then let \(U_{I, (\rho_i)} = \{(A_{0, b}, U) \in V : b \in B_{I, (\rho_i)}\}\).
\begin{lemma}
    \(\mathbb{C}^m = \bigcup_{I, (\rho_i)} B_{I, (\rho_i)}\), where \(I\) ranges over all subsets of \(\{1, ..., k\}\), and \((\rho_i)\) ranges over all maps \(I \to \mathbb{N}_{>0}\) such that \(\rho_i\) and \(\lambda_i - \rho_i\) are both decreasing with \(i\).
    TODO: something about overlap being small (they're not disjoint).  
    They are distinct.
    Is that good enough for my purposes?
    Ah, I think it is good enough if I instead say that the sets of \(b\)'s are distinct.
\end{lemma}
\begin{proof}
    Let \(b \in \mathbb{C}^m\).
    Let \(I' = \{(p_i, q_i) : p_i = n > q_i \land \max_{j : q_j = q_i} p_j > \max_{j : q_j < q_i} p_j\}\).
    Then let \(I = \{\max\{i : (p_i, q_i) = (p, q)\} : (p, q) \in I'\}\).
    Choosing the maximal \(i\) is an arbitrary choice; we just have to pick one.
    For \(i \in I\), \(\rho_i := p_i\).
    All we need to do is show that each \(\rho_i\) is positive, and that \(\rho_i\) and \(\lambda_i - \rho_i\) are decreasing.
    Then we get that \((A_{0,b}, U) \in U_{I, (\rho_i)}\), and we are done.
    \par Recall that by convention the maximum of the empty set is zero.
    So, the fact that \(p_i > \max_{j > i} p_j\) tells us that \(\rho_i = p_i\) is positive.
    \par Let \(i, j \in I\) with \(i < j\).
    Since \(\lambda_i \geq \lambda_j\), we have either \(p_i \geq p_j\) or \(q_i \geq q_j\).
    Further, since \((p_i, q_i) \neq (p_j, q_j)\) (clear from definition of \(I\)), either \(p_i > p_j\) or \(q_i > q_j\).
    And for \(i, j \in I\) we clearly have \(p_i > p_j \iff q_i > q_j\).
    So in either case, both \(p_i > p_j\) and \(q_i > q_j\), and consequently both \(\rho_i > \rho_j\) and \(\lambda_i - \rho_i > \lambda_j - \rho_j\).
\end{proof}

@@

\begin{corollary}\label{bs_iso}
    \[B_{I, (\rho_i)} \cong \prod_{i \in I} \mathbb{C}^{blah} \times \prod_{i \notin I}\]
\end{corollary}

@@

\begin{lemma}
    \(U_{I, (\rho_i)} \cong B_{I, (\rho_i)} \times X_\mu\).
\end{lemma}
\begin{proof}
    For \(b \in B_{I, (\rho_i)}\), let \(P_b\) be the change-of-basis matrix, with columns given by the Jordan basis of the previous lemma, so that \(J(\mu) = P_b^{-1} A_{0,b} P_b\).
    From looking at the Jordan basis of the previous lemma (the basis can be expressed in terms of \(A, \lambda, e_{ij}, f_i, I, \rho_i, b\)), it is clear that the map \(P : B_{I, (\rho_i)} \to \GL_{m + n}\), given by \(b \mapsto P_b\), is algebraic.
    \par Now, we remark that the Springer fiber at \(A_{0,b}\) is simply \(\{P_b U : U \in X_\mu\}\).
    This gives us the isomorphism \(B_{I, (\rho_i)} \times X_\mu \to U_{I, (\rho_i)}\).
    \[(b, U) \mapsto (P_b J(\mu) P_b^{-1}, P_b U),\]
    with inverse
    \[(A_{0,b}, U) \mapsto (b, P_b^{-1} U).\]
\end{proof}
At last we have reached the base of our recursive procedure, and we begin to propagate upwards.
We write \((X_{\mu, \alpha})_\alpha\) to denote the irreducible components of \(X_\mu\).
\begin{lemma}
    The irreducible components of \(U_{I, (\rho_i)}\) are exactly the subvarieties of the form
    \[\{(P_b J(\mu) P_b^{-1}, P_b U) : b \in B_{I, (\rho_i)}, U \in X_{\mu, \alpha}\}.\]
    There are blah of these, and each has dimension blah.
\end{lemma}
\begin{proof}
    Components of \(B_{I, (\rho_i)} \times X_\mu\) are obviously blah.
    Their dimensions are blah, and there are... of them.
    taking image under the isomorphism gives the desired result.
\end{proof}
From \cref{bs_iso}, we know that \(B_{I, (\rho_i)}\) is irreducible, of dimension blah.

The irreducible components of \(U_{I, (\rho_i)}\) are blah.
Thus, the irreducible components of \(U_0\) are blah.

>>>>

\section{Finding the irreducible components of \(U_0\)}
\subsection{Introduction}
Let \(X\) be nilpotent with Jordan basis \((e_{ij}')_{1 \leq i \leq k, 1 \leq j \leq \lambda_i'}\).
Let \(V = \{(A_{a,b}, U) : \forall i. \; A_{a,b} U_{i + 1} \subseteq U\}\).
Let
\[U_0 = \{(A_{0,b}, U) \in V\}.\]
We will find the irreducible components of \(U_0\).
\par We write \(b_{ij}\) to denote the projection of \(b \in \mathbb{C}^m\) onto \(e_{ij}\).
For each row \(i\), let \(p_i(b) = \max\{j : b_{ij} \neq 0\}\) (the maximum of the empty set is zero).
Then set \(q_i(b) = \lambda_i - p_i(b)\).
When it is clear enough from context, we will just write \(p_i\) and \(q_i\) instead of \(p_i(b)\) and \(q_i(b)\).
\par For any \(I \subseteq \{1, ..., k\}\) and any values \((\rho_i)_{i \in I}\) %such that (1) \(\rho_i \leq \lambda_i\), (2) \(\rho_i\) is decreasing with \(i\), and (3) \(\lambda_i - \rho_i\) is decreasing with \(i\)
, define 
\[B_{I, (\rho_i)} := \{b \in \mathbb{C}^m : \]
\[\{(p_i, q_i) : i \in I\} = \{(p_i, q_i) : n > q_i \land p_i = \max_{j : q_j = q_i} p_j > \max_{j : q_j < q_i} p_j \} \] 
\[\land \forall i \in I. \; p_i = \rho_i \}.\]
Then let \(U_{I, (\rho_i)} = \{(A_{0, b}, U) \in V : b \in B_{I, (\rho_i)}\}\).
\begin{lemma}
    \(\mathbb{C}^m = \bigcup_{I, (\rho_i)} B_{I, (\rho_i)}\), where \(I\) ranges over all subsets of \(\{1, ..., k\}\), and \((\rho_i)\) ranges over all maps \(I \to \mathbb{N}_{>0}\) such that \(\rho_i\) and \(\lambda_i - \rho_i\) are both decreasing with \(i\).
    TODO: something about overlap being small (they're not disjoint).  
    They are distinct.
    Is that good enough for my purposes?
    Ah, I think it is good enough if I instead say that the sets of \(b\)'s are distinct.
\end{lemma}
\begin{proof}
    Let \(b \in \mathbb{C}^m\).
    Let \(I' = \{(p_i, q_i) : p_i = n > q_i \land \max_{j : q_j = q_i} p_j > \max_{j : q_j < q_i} p_j\}\).
    Then let \(I = \{\max\{i : (p_i, q_i) = (p, q)\} : (p, q) \in I'\}\).
    Choosing the maximal \(i\) is an arbitrary choice; we just have to pick one.
    For \(i \in I\), \(\rho_i := p_i\).
    All we need to do is show that each \(\rho_i\) is positive, and that \(\rho_i\) and \(\lambda_i - \rho_i\) are decreasing.
    Then we get that \((A_{0,b}, U) \in U_{I, (\rho_i)}\), and we are done.
    \par Recall that by convention the maximum of the empty set is zero.
    So, the fact that \(p_i > \max_{j > i} p_j\) tells us that \(\rho_i = p_i\) is positive.
    \par Let \(i, j \in I\) with \(i < j\).
    Since \(\lambda_i \geq \lambda_j\), we have either \(p_i \geq p_j\) or \(q_i \geq q_j\).
    Further, since \((p_i, q_i) \neq (p_j, q_j)\) (clear from definition of \(I\)), either \(p_i > p_j\) or \(q_i > q_j\).
    And for \(i, j \in I\) we clearly have \(p_i > p_j \iff q_i > q_j\).
    So in either case, both \(p_i > p_j\) and \(q_i > q_j\), and consequently both \(\rho_i > \rho_j\) and \(\lambda_i - \rho_i > \lambda_j - \rho_j\).
\end{proof}

@@

\begin{corollary}\label{bs_iso}
    \[B_{I, (\rho_i)} \cong \prod_{i \in I} \mathbb{C}^{blah} \times \prod_{i \notin I}\]
\end{corollary}

\subsection{Irreducible components of \(U_{I, (\rho_i)}\)}

@@

\begin{lemma}
    \(U_{I, (\rho_i)} \cong B_{I, (\rho_i)} \times X_\mu\).
\end{lemma}
\begin{proof}
    For \(b \in B_{I, (\rho_i)}\), let \(P_b\) be the change-of-basis matrix, with columns given by the Jordan basis of the previous lemma, so that \(J(\mu) = P_b^{-1} A_{0,b} P_b\).
    From looking at the Jordan basis of the previous lemma (the basis can be expressed in terms of \(A, \lambda, e_{ij}, f_i, I, \rho_i, b\)), it is clear that the map \(P : B_{I, (\rho_i)} \to \GL_{m + n}\), given by \(b \mapsto P_b\), is algebraic.
    \par Now, we remark that the Springer fiber at \(A_{0,b}\) is simply \(\{P_b U : U \in X_\mu\}\).
    This gives us the isomorphism \(B_{I, (\rho_i)} \times X_\mu \to U_{I, (\rho_i)}\).
    \[(b, U) \mapsto (P_b J(\mu) P_b^{-1}, P_b U),\]
    with inverse
    \[(A_{0,b}, U) \mapsto (b, P_b^{-1} U).\]
\end{proof}
At last we have reached the base of our recursive procedure, and we begin to propagate upwards.
We write \((X_{\mu, \alpha})_\alpha\) to denote the irreducible components of \(X_\mu\).
\begin{lemma}
    The irreducible components of \(U_{I, (\rho_i)}\) are exactly the subvarieties 
    \[U_{I, (\rho_i), \alpha} := \{(P_b J(\mu) P_b^{-1}, P_b U) : b \in B_{I, (\rho_i)}, U \in X_{\mu, \alpha}\}.\]
    These are distinct.  
    There are blah of them, and each has dimension blah.
\end{lemma}
\begin{proof}
    We know from \cref{bs_iso} that \(B_{I, (\rho_i)}\) is irreducible.
    So, the components of \(B_{I, (\rho_i)} \times X_\mu\) are obviously \(B_{I, (\rho_i)} \times X_{\mu, \alpha}\).
    Their dimensions are blah, and there are... of them.
    taking image under the isomorphism gives the desired result.
\end{proof}

\subsection{Conclusion}
\begin{lemma}
    The irreducible components of \(U_0\) are exactly the \(U_{I, (\rho_i), \alpha}\).
    They are distinct.
    There are blah of them, each of dimension blah.
\end{lemma}
